Analysis of the inference of ratings and rankings on Higher Order Networks with complex topologies

### 2025-09-03 draft/draft.tex
progress

The dataset covers five soccer leagues: England, France, Germany, Italy, and Spain.
Each league consists of 20 teams, except for Germany, which has 18.
Every pair of teams $i,j$ plays two matches: one with $i$ as the home team and $j$ as the away team, and in the other match it is the other way around.
The teams are indexed or ranked according to their position in that year’s final standings.
The standings is characterized by some tournament points $\pi_i$ for each team $i$ in each league.
Therefore, the condition $i<j$ holds if and only if $\pi_i>\pi_j$ for any two teams $i$ and $j$ in a given league.

For each pair of teams $i$ and $j$ in a given league, several pairwise metrics $m_{ij}\in \mathbb{R}$ are provided.
Each metric takes a positive value if it presumably favors team $i$ over team $j$, and a negative value otherwise.
The different types of empirically determined metrics being considered are:
\begin{enumerate}
\item {\bf crossings:}
\item {\bf counterattacks:}
\item {\bf pressure loss:}
\item {\bf build-up time:}
\item {\bf direct play:}
\item {\bf pressure points:}
\item {\bf shots:}
\item {\bf flow rate:}
\item {\bf maintenance time:}
\item {\bf middle-zone time:} bla bla.
\end{enumerate}
Namely, $m_{ij}$ corresponds to $i$ being the home team and $j$ being the away team.
Additionally, metric-styles derived from the Singular Value Decomposition of previous metrics are also considered~\cite{chacoma2025data}.
Moreover, in order to test the inference of ratings, two types of ground-truth metrics are also defined.
Namely, the rank-based metric $m_{ij}=\frac{1}{2}(i-j)$ and the tournament point metric $m_{ij}=\frac{1}{2}(\sigma_j-\sigma_i)$.
Notice, $m_{ij}$ is generally independent of the metric $m_{ji}$ since they belong to different matches between $i$ and $j$.
In the case of the ground-truth metrics, it holds that $m_{ij}=-m_{ji}$.

{\color{red}[HERE...]}

Each matrix $\phi$ of scores $\phi_{ij}$ defines a fully connected undirected network for the teams in a league.
In some experiments, some fraction of links is removed from the network according to some criterium, defining a graph $\mathcal{G}$ that is not fully connected.
The graph is promoted to a simplicial complex of 0-simplices, 1-simplices and 2-simplices composed by the nodes, the links and the triplets of connected nodes in $\mathcal{G}$, respectively.
A triplet $ijk$ of the nodes $i$, $j$ and $k$ of $\mathcal{G}$ is connected if the links $ij$, $ik$ and $jk$ belong to $\mathcal{G}$.

>>>>

The dataset covers five soccer leagues: England, France, Germany, Italy, and Spain.
Each league consists of $N=20$ teams, except for Germany, which has $N=18$.
Every pair of teams $i,j$ plays two matches: one with $i$ as the home team and $j$ as the away team, and in the other match it is the other way around.
The teams are indexed or ranked according to their position in that year’s final standings.
The standings is characterized by some tournament points $\pi_i$ for each team $i$ in each league.
Therefore, the condition $i<j$ holds if and only if $\pi_i>\pi_j$ for any two teams $i$ and $j$ in a given league.

For each pair of teams $i$ and $j$ in a given league, several pairwise metrics $m_{ij}\in \mathbb{R}$ are provided.
Each metric takes a positive value if it presumably favors team $i$ over team $j$, and a negative value otherwise.
The different types of empirically determined metrics being considered are:
\begin{enumerate}
\item {\bf crossings:}
\item {\bf counterattacks:}
\item {\bf pressure loss:}
\item {\bf build-up time:}
\item {\bf direct play:}
\item {\bf pressure points:}
\item {\bf shots:}
\item {\bf flow rate:}
\item {\bf maintenance time:}
\item {\bf middle-zone time:} bla bla.
\end{enumerate}
Namely, $m_{ij}$ corresponds to $i$ being the home team and $j$ being the away team.
Additionally, metric-styles derived from the Singular Value Decomposition of previous metrics are also considered~\cite{chacoma2025data}.
Notice, $m_{ij}$ is generally independent of the metric $m_{ji}$ since they belong to different matches between $i$ and $j$.
As a consequence, $m_{ij}+m_{ji}$ is generally different from zero and is the net contribution between teams $i$ and $j$.

With each metric, a 1-cochain is defined as
$$
f_{ij}
=
\frac{1}{2}(m_{ij}+m_{ji})
$$
Moreover, in order to test the inference method, two types of ground-truth 1-cochains are also defined.
Namely, the rank-based 1-cochain $f_{ij}=\frac{1}{2}(i-j)$ and the tournament point 1-cochain $f_{ij}=\frac{1}{2}(\pi_j-\pi_i)$.
Note, $f_{ij}=-f_{ji}$ holds by definition.
These 1-cochains $f$ define a fully connected network or graph, since all pairs of teams play two matches against each other in each league.
Each of these fully connected graphs of $N=n+1$ nodes is promoted to a simplicial complex compose of 0-simplices labeled by $i$, 1-simplices labeled by $ij$ and 2-simplices labeled by $ijk$ where $i,j,k$ are different elements in $\{0,1,2,...,n\}$.
